---
output: 
  md_document:
    fig_height: 0.3
    fig_width: 3
    variant: markdown_github
---

```{r options, echo=FALSE}
knitr::opts_chunk$set(comment="#", tidy=FALSE)
```
# chroma

`chroma` is a package for parsing and formating colors in various specifications, manipulating colors, and creating nice color scales and palettes.

## Installation

`chroma` is not on CRAN yet. To get it from this page, install the package `devtools` and

```{r install, eval=F}
devtools::install_github("jiho/chroma")
library("chroma")
```
```{r load, echo=F}
devtools::load_all(".", quiet=TRUE)
```

## Color parsing

Parse colors in various specifications

```{r parse}
rgb(r=0.5, g=0.5, b=0.5)
hsv(h=120, s=0.5, v=0.5)
hsl(h=120, s=0.5, l=0.5)
hsi(h=120, s=0.5, i=0.5)
hcl(h=120, c=0.5, l=0.5)
lab(l=0.5, a=-0.5, b=0.2)
hex("#393939")
hex("F39")
css("rgb(100,100,100)")
temperature(5600)
```

Parse directly from a matrix

```{r parse_matrix, comment="#"}
x <- matrix(c(0.2, 0.5, 0.5, 0.5, 0.6, 0.4), ncol=3)
print(x)
rgb(x)
```

## Color formating

Convert back in various formats

```{r convert}
as.rgb("coral1")
as.hsv("coral1")
as.hsl("coral1")
as.hsi("coral1")
as.hcl("coral1")
as.lab("coral1")
as.hex("coral1")
as.css("coral1")
as.temperature("coral1")
```

## Color manipulation

Slightly modify a base color

```{r modify_color}
col <- "#7BBBFE"
show_col(c(col, brighten(col), darken(col)))
show_col(c(col, desaturate(col), saturate(col)))
```

Make a color semi-transparent

```{r alpha}
col <- "#7BBBFE"
show_col(c(col, alpha(col), alpha(col, 0.2)))
```

Mix two colors

```{r mix}
show_col(c("#7BBBFE", "#FDFF68", mix("#7BBBFE", "#FDFF68")))
```

Compute the contrast between two colors

```{r contrast}
contrast("darkblue", "darkgreen")
contrast("yellow", "darkgreen")
```

Extract or set a color channel

```{r channel}
col <- "#7BBBFE"
channel(col, model="hcl", "h")
channel(col, model="hsv", "s")
channel(col, model="hsi", "i")
channel(col, model="rgb", "r")

col1 <- col2 <- col
channel(col1, model="hcl", "h") <- 120
channel(col2, model="hcl", "l") <- 1
show_col(c(col, col1, col2))
```

Compute or set the perceived luminance of a color

```{r luminance}
luminance(c("red", "yellow", "darkblue"))

col1 <- col2 <- col <- "#7BBBFE"
luminance(col)
luminance(col1) <- 0.6
luminance(col2) <- 0.2
show_col(c(col, col1, col2))
```

## Color scales and palettes

All scales and palettes are organised the same way:

- functions ending in `*_scale` return a *function* that takes a numeric vector `x` as argument and returns the corresponding colors in the scale.
- functions ending in `*_map` are shortcuts that build the scale, map the values, and return the colors.
- functions ending in `*_palette` return a *function* that takes an integer `n` as argument and returns `n` equally spaced colors along the scale.
- functions ending in `*_colors` (or `*.colors`) are shortcut that create the palette and return the `n` colors.

```{r scales}
x <- 0:10/10
s <- color_scale()
s(x)
# or
color_map(x)

n <- 11
p <- color_palette()
p(n)
# or
interp_colors(n)
```



Palettes can be built by interpolating between colors

```{r scale_interp}
show_col(interp_colors(10))
show_col(interp_colors(10, colors=c("#2D2B63", "#F7FF84")))
show_col(interp_colors(10, colors=c("#2D2B63", "#FB3C44", "#F7FF84")))
show_col(interp_colors(10, colors=c("#2D2B63", "#FB3C44", "#F7FF84"), interp="bezier"))
```

Preset palettes are available, from colorbrewer

```{r colorbrewer, fig.height=0.3*35, fig.width=2.1}
show_col(lapply(brewer_info$name, function(x) {brewer.colors(n=7, name=x)}))
```

Or viridis

```{r viridis, fig.height=0.3*4}
show_col(
  viridis_colors(10),
  magma_colors(10),
  plasma_colors(10),
  inferno_colors(10)
)
```

Or cubehelix

```{r cubehelix, fig.height=0.3*4}
show_col(
  cubehelix_colors(10),
  cubehelix_colors(10, h=300, rot=-0.75),
  cubehelix_colors(10, h=120, rot=0.5),
  cubehelix_colors(10, h=300, rot=0.5)
)
```

New, perceptually appropriate, palettes can be built in HCL space, for either discrete

```{r hue}
show_col(hue_colors(10))
```

or continuous variables

```{r chroma_light}
show_col(chroma_colors(10))
show_col(light_colors(10))
```

---

Happy coloring!
